---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Interaction Graphs package 'integr' 

The Interaction Graphs package 'integr' is an implementation of Interaction Analysis of Aleks Jakulin (http://stat.columbia.edu/~jakulin/Int/). It provides a GraphViz graph of 3-way interactions (i.e. conditional Information Gain) based on a provided discrete data.frame (i.e. all columns are factors). This graph serves to analyse the correlations between the attributes, as well to select the most informative attributes in the context of predictive machine learning modeling. The obtained interaction graph can be exported to various output formats: Graphviz file, SVG, PNG, PDF, or PS images.

In the following text, installation guides (from CRAN and GitHub) and a simple example for using the package are provided. For more detailed information on Interaction Graphs methodology, please consult the Tutorial vignette of this package using the command, after the package is installed: 

```{r eval=FALSE}
vignette("IntGraphTutorial", package = "integr")
```
or visit the webpage of Aleks Jakulin (http://stat.columbia.edu/~jakulin/Int/).

## Installation

One can install the released version of integr from [CRAN](https://CRAN.R-project.org) with:

```{r eval = FALSE}
install.packages("integr")
```

Or from GitHub with:
```{r eval = FALSE}
library("devtools") #devtools package needs to be installed first!
install_github("peleplay/integr", build_opts = c("--no-resave-data", "--no-manual"), build_vignettes = TRUE)
```

## Example

A basic example which shows how to generate an Interaction Graph and export it to various formats is provided below. The same example is used in the Tutorial vignette.

```{r, eval=FALSE}

#load integr package (needs to be installed first!)
library("integr")

#read Golf toy-dataset
data("golf")

#create an Interaction graph object
g <- interactionGraph(golf, classAtt = "Play", intNo = 10, speedUp = FALSE)

#plot the Interaction graph object (in RStudio!)
plotIntGraph(g)

#export the Interaction graph object to a Graphviz file
igToGrViz(g, path = "myFolder", fName = "myInteractionGraph")

#export the Interaction graph object to a SVG image
igToSVG(g, path = "myFolder", fName = "myInteractionGraph", h = 2000)

#export the Interaction graph object to a PNG image
igToPNG(g, path = "myFolder", fName = "myInteractionGraph", h = 2000)

#export the Interaction graph object to a PDF image
igToPDF(g, path = "myFolder", fName = "myInteractionGraph", h = 2000)

#export an Interaction graph object to a PS image
igToPS(g, path = "myFolder", fName = "myInteractionGraph", h = 2000)
```

_path_ parameter is a string indicating the path (folder) in which the output should be saved. If not specified differently, working directory by default;

_fName_ parameter is a string indicating the name of the output. It should be defined without extension and without spaces. If not specified differently, 'InteractionGraph' by default;

_h_ is the desired height of the output image in pixels. If not defined differenlty, 2000 by default.
